---
title: 'Using fooPackage'
author: "Jose Antonio Garcia Ramirez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The _fooPackage_ was built and designed to handle a dataset whose names are of the form: *accident_year.csv* where _year_ is a character indicating a numbre for example the following names 

- *accident_2013.csv*
- *accident_2014.csv*

This is because it was thought to exploit the following dataset which can be downloaded from [url](https://d3c33hcgiwev3.cloudfront.net/_e1adac2a5f05192dc8780f3944feec13_fars_data.zip?Expires=1487203200&Signature=jpCri1RCtoJ1S8Y5OevP7xzs9gEhY2yEYWhVLdtH--pP7u33Pgc2E8g1T52qgB-nMFoIg2fRS5do0oqCcef~VYkqK0h1qiO9k7tlLSDGMd~Ob~XCMMd8IMvrPE56YbhdBQm4UsPjqkshkC7IMtQVGHwQz4kvycAn-~VIogHfCU0_&Key-Pair-Id=APKAJLTNE6QMUY6HBC5A) and will be used to exemplify the use of package (the data is whithin the fooPackages)

The first function to exemplify is _make_filename_ which construct a string with the name of the file of interest.

```{r}
file.name <- make_filename(2013)  
file.name
fars_read(file.name)
```





## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
