---
title: "A concise visualization of earthquakes in the world"
author: "J. Antonio García, jose.ramirez@cimat.mx"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE, cache=TRUE}
knitr::opts_chunk$set( 
  collapse = TRUE,
  comment = "#>", 
  cache = TRUE 
)
```

```{r libraries} 
library(FooR)
library(xtable)
library(pander)
library(dplyr)
library(ggplot2)
```

This vignette two process:
  - Cleaning 
  - Two different visualizations (over time - like [Doctor Who](https://en.wikipedia.org/wiki/Doctor_Who) would see it - and country)
  
  About the NOAA earthquake data. the data come can be obtained from [NOAA Significant Earthquake Database](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1).

## Package content

There are ten functions within the package:

- `fecha()`
- `eq_clean_data()`
- `eq_location_clean()`
- `theme_timeline()`
- `GeomTimeline()`
- `GeomTimelineLabel()`
- `theme_timeline()`
- `geom_timeline()`
- `geom_timeline_label()`
- `eq_map()`

But only the following functions are exported:

- `fecha()`
- `eq_clean_data()`
- `eq_location_clean()`
- `theme_timeline()`
- `GeomTimeline()`
- `GeomTimelineLabel()`
- `theme_timeline()`
- `eq_map()`


## Data
The package contains the dataset obtained from [National Geophysical Data Center / World Data Service (NGDC/WDS): Significant Earthquake Database. National Geophysical Data Center, NOAA. doi:10.7289/V5TD9V7K](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1) consulted on the date "`r format(Sys.time(), '%d %B, %Y')`" and and obviously the dataset is called *data*


```{r print1, echo = FALSE, }
NOAA <- data
```

 
 -------------------------------------------------------------------
  I_D   FLAG_TSUNAMI   YEAR    MONTH   DAY   HOUR   MINUTE   SECOND 
 ----- -------------- ------- ------- ----- ------ -------- --------
   1         NA        -2150    NA     NA     NA      NA            
 
   3         NA        -2000    NA     NA     NA      NA       NA   
 
 --------------------------------------------------------------------------
  FOCAL_DEPTH   EQ_PRIMARY   EQ_MAG_MW   EQ_MAG_MS   EQ_MAG_MB   EQ_MAG_ML 
 ------------- ------------ ----------- ----------- ----------- -----------
      NA           7.3          NA          NA          NA          NA     
 
      18           7.1          NA          7.1         NA          NA     
 --------------------------------------------------------------------------
 
  
 ------------------------------------------------------------
  EQ_MAG_MFA   EQ_MAG_UNK   INTENSITY     COUNTRY      STATE 
 ------------ ------------ ----------- -------------- -------
      NA          7.3          NA          JORDAN       NA   
 
      NA           NA          10       TURKMENISTAN    NA   
 ------------------------------------------------------------
 
 
  
 ------------------------------
         LOCATION_NAME         
 ------------------------------
  JORDAN: BAB-A-DARAA,AL-KARAK 
 
        TURKMENISTAN: W        
 ------------------------------

## Data cleaning

It is evident that the data set has numerous columns with  [null](https://en.wikipedia.org/wiki/Null_(SQL)) or missing values, <span style="color:RED"> although a future and interesting work for this package would be to apply data imputation techniques to the dataset </span>. For the above you can use the function *eq_clean_data()* and *eq_location_clean*

```{r eq_clean_data_eq_location_clean, fig.width = 7, fig.height = 7, }
NOAA <- eq_clean_data(NOAA)
NOAA <- eq_location_clean(NOAA)
```

## Visualization of earthquakes by time and place :

  1. [Doctor-Who](https://en.wikipedia.org/wiki/Doctor_Who)'s [view](https://en.oxforddictionaries.com/definition/view) ( or style1),          *_all the time_* and _*all the countries*_ at once.
     You can visualize earthquakes in all dataset's time variable *YEAR* without 
     filter by the columns *COUNTRY* and *YEAR* and using the package *ggplot* and        the recently implemented *geom_*, *geom_timeline*  and *geom_timeline_label*.        (you must pay attention with the column so that the visualization is of      a       suitable size and not empty, as in every ggplot´s function, one must have            congruent with the variables' name of the dataset  and the parameters of             *ggplot2::aes* function.)[^1] 

```{r doctor,  fig.width = 16, fig.height = 17}
NOAA %>%
filter( DEATHS>0) %>%
     ggplot(aes(x = Date,
                y = COUNTRY,
                color = DEATHS,
                size = as.numeric(EQ_PRIMARY)
     )) +
     geom_timeline() +
     #geom_timeline_label(aes(label = LOCATION_NAME)) +
     theme_timeline() +
     labs(size = "Richter scale value", color = "# deaths",
          title = "Figure 1 : Doctor Who's view")
```


  2. No-Doctor-Who's (style2), not all the time neither all the countries at once.
     You can visualize earthquakes in a single country after 
     filter by the column *COUNTRY* or many others, (you must pay       attention with the columns of _YEAR_ and _DEATHS_ so that the visualization is       of a suitable size and not empty, as in every ggplot´s function, one must have      congruent with the variables' name of the dataset  and the parameters of             *ggplot2::aes* function.) 
     
```{r human, fig.width = 12, fig.height = 3}
NOAA %>%
filter(COUNTRY %in% c("USA"), 2020 > YEAR,  YEAR > 2000, DEATHS>0) %>%
     ggplot(aes(x = Date,
                y = COUNTRY,
                color = DEATHS,
                size = as.numeric(EQ_PRIMARY)
     )) +
     geom_timeline() +
     geom_timeline_label(aes(label = LOCATION_NAME)) +
     theme_timeline() +
     labs(size = "Richter scale value", color = "# deaths",
     title = "Figure 2: Human vision")
```

  3. Style3, not all the time neither all the countries at once.
     You can visualize earthquakes in a single country after 
     filter by the column *COUNTRY* or many others, (you must pay 
     attention with the columns of _YEAR_ and _DEATHS_ as in style2)
     
```{r human2, fig.width = 12, fig.height = 6}
NOAA %>%
filter(COUNTRY %in% c("USA", "MEXICO", "JAPAN"), 2020 > YEAR,  YEAR > 2000, DEATHS>0) %>%
     ggplot(aes(x = Date,
                y = COUNTRY,
                color = DEATHS,
                size = as.numeric(EQ_PRIMARY)
     )) +
     geom_timeline() +
     geom_timeline_label(aes(label = LOCATION_NAME)) +
     theme_timeline() +
     labs(size = "Richter scale value", color = "# deaths",
     title = "Figure 3: Human vision")
```

# Visualize earthquakes on a pretty map

Using the *leaflet* package it is very easy to make beautiful interactive maps using the function *eq_map()*.There are two options to use this function one without using the function *eq_create_label()* from where we get the following map:

```{r mapa_simple, fig.width = 8, fig.height = 6}
NOAA %>% 
   dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(Date) >= 2000) %>% 
  eq_map(annot_col = "DATE")
```

And the other one using it

```{r mapa_con_marcas, fig.width = 8, fig.height = 6}
NOAA %>%
 dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(Date) >= 2000) %>% 
  dplyr::mutate(popup_text = eq_create_label(.)) %>% 
  eq_map(annot_col = "popup_text")
```

[^1]: Thanks to my [brother](https://www.facebook.com/lucaridavidmisticmaster) and my friend [Claudia](https://www.facebook.com/Ayla.occord?pnref=friends.search) for the idea about figure 1.
